---
title: "Poisson Chapter Three Feb 10 2020"
author: "Victoria_Field"
date: "2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(ggplot2)
library(pscl)
library(lme4)
library(GLMMadaptive)
library(glmmTMB)
library(bbmle)
library(DHARMa)
```

```{r}
#import sum.data.edited 
count<-read.csv("sum.data.1.2.csv", header=TRUE, colClasses = c("numeric","factor","factor","factor", rep("numeric",7)))
count$WEEKS_IN<-round(count$DAYS_IN/7,0)
count$MONTHS_IN<-round(count$DAYS_IN/30,0)
count$MONTHS_IN<-as.factor(count$MONTHS_IN)

#re-order factor level for 'LAKE_INV.YEAR
count$LAKE_INV.YEAR <- factor(count$LAKE_INV.YEAR, levels = c("Crooked - Not Invaded", "Tully - Not Invaded", "Cazenovia_A - 1997", "Cazenovia_B - 1997", "Owasco - 1997", "Honeoye - 1998", "EatonBrook - 1999", "Silver - 2001", "Craine - 2007", "UpperLittleYork_A - 2013", "Song - 2017"))
```

#Distribution of the Abundance Count 

```{r}
summary(count$Count)
plot(count$Count~count$WEEKS_IN)
hist(count$Count, breaks=10)
hist(log(count$Count))
```

#Poisson Model for Abundance (from: https://online.stat.psu.edu/stat504/node/169/)
```{r}
model<-glmer(Count ~ DAYS_IN + Years_Invaded + Dreissenids + (1|LAKE), family=poisson(link=log), data=count)
summary(model)
```

```{r}
model2<-glmer(Count ~ WEEKS_IN  + (1|LAKE), family=poisson(link=log), data=count)
summary(model2)
```

You can also get the predicted count for each observation and the linear predictor values from R output by using specific statements such as:

```{r}
#to get the predicted count for each observation: 
#e.g. for the first observation E(y1)=3.810

print<-data.frame(count, pred=model$fitted)
print

#note the linear predictor values

model$linear.predictors
exp(model$linear.predictors)
```

Interpretations: 
-Standard error of beta=0.045 is 0.0027

-Slope of beta is statistically significant with z-value=16.57 and a p-value of <2e-16

-We can also see that although the predictor is significant the model does not fit well. Given the value of the residual deviance statistic of 4296.9 with 49 df, the p-value is zero and the Value/DF=4296.9/49=87.69 is much bigger than 1, so the model does not fit well. The lack of fit maybe due to missing data, covariates or overdispersion.

  -Let us assume for now that we do not have any other covariates, and try to adjust for overdispersion to see if we can improve the model fit.
  
##Change the Model: Adjusting for Overdispersion

In the above model we detect a potential problem with overdispersion since the scale factor, e.g., Value/DF for the residual deviance/df, is much greater than 1. 

```{r}
#First, look at sample means and variance 
tapply(count$Count, count$MONTHS_IN, function(x)c(mean=mean(x),variance=var(x)))
```

```{r}
#Estimate dispersion factor using quasipoisson family 

model.disp<-glm(count$Count~count$MONTHS_IN, family=quasipoisson(link=log), data=count)
summary.glm(model.disp)
summary.glm(model.disp)$dispersion
```

```{r}
#Estimate the dispersion factor using negative binomial 

nb.fit<-glm(count$Count~count$MONTHS_IN, data=count, family=negative.binomial(theta=1, link="identity"), start=model$coef)
summary(nb.fit)

```

What could be another reason for poor fit besides overdispersion? How about missing other explanatory variables? Can we improve the fit by adding other variables?

##Change the Model: Include ‘LAKE’ as a Qualitative Predictor

#GLM with poisson family (from:https://www.statmethods.net/advstats/glm.html)

```{r}
fit<-glm(Count ~ WEEKS_IN + Years_Invaded, data=count, family=poisson(link="log"))
summary(fit)
```

#Zero-inflated Poisson regression (from https://stats.idre.ucla.edu/r/dae/zinb/)

```{r}
m1 <- zeroinfl(Count ~ WEEKS_IN | Dreissenids,
  data = count, dist = "negbin", EM = TRUE)
summary(m1)
```

#Zero-inflated Poisson regression (from https://www.r-bloggers.com/zero-inflated-poisson-and-negative-binomial-models-with-glmmadaptive/)

```{r}
zi.mm<-mixed_model(Count ~ WEEKS_IN, random = ~ 1 | LAKE, data = count,
                   family = poisson())
zi.mm
```

#Zero-Inflated Poisson MEM 

```{r}
fit_zipoisson <- glmmTMB(Count ~ WEEKS_IN + Years_Invaded + (1|LAKE), data=count, ziformula= ~ Dreissenids, family=poisson(link="log"))
summary(fit_zipoisson)
```

```{r}
fit_zinbinom <- update(fit_zipoisson,family=nbinom2)
summary(fit_zinbinom)

fit_zinbinom1 <- update(fit_zipoisson,family=nbinom1)
summary(fit_zinbinom1)

AICtab(fit_zipoisson,fit_zinbinom)
```

```{r}
ziresiduals<-residuals(fit_zinbinom, type = c("response", "pearson"))
plot(ziresiduals)
abline(0,0)
```

##Model Diagnostics
```{r}
zires<-simulateResiduals(fit_zinbinom)
plot(zires)

testResiduals(zires)
```

##No zero-inflated 

```{r}
fit_poisson <- glmmTMB(Count ~ WEEKS_IN + Years_Invaded + (1|LAKE), data=count, family=poisson(link="log"))
summary(fit_poisson)

fit_nbinom <- update(fit_poisson,family=nbinom2)
summary(fit_nbinom)

fit_nbinom1 <- update(fit_poisson,family=nbinom1)
summary(fit_nbinom1)

AICtab(fit_poisson,fit_nbinom, fit_nbinom1)
```

```{r}
fixef(fit_nbinom) #pulls fixed effect coefficients

exp(0.07973) # equals 1.083; increasing `WEEKS_IN` by 1 multiplies the mean value of `Count` by 1.083
exp(0.25161) # equals 1.286; increasing `YEARS_INVADED` by 1 multiplies the mean value of `Count` by 1.286
```

##Model Diagnostics
```{r}
residuals<-residuals(fit_nbinom, type = c("response", "pearson"))
plot(residuals)
abline(0,0)
hist(residuals)
```

```{r}
res<-simulateResiduals(fit_nbinom)
plot(res)

testResiduals(res)

testZeroInflation(res)
```

```{r}
summary(lm(count$Abundance_m2 ~ count$WEEKS_IN))
```